---
title: "HW_Lab 1 TSA"
author: "Maria Isaac"
output: 
  pdf_document: 
    latex_engine: lualatex
---

```{r}
# necessary library
library(astsa)
```

```{r}
# Function to generate and plot white noise, ACF, and autocovariance
analyze_white_noise <- function(n) {
  # Generating white noise
  wn <- rnorm(n)
  
  # plotting layout (2 plots: trajectory and correlogram)
  par(mfrow=c(2,1))
  
  # Plotting white noise trajectory
  plot(wn, type='l', ylab="White Noise", main=paste("Trajectory of White Noise (n =", n, ")"))
  points(wn)
  
  # Plotting ACF (correlogram)
  acf(wn, lag.max=10, plot=TRUE, main=paste("Correlogram (n =", n, ")"))
  
  # Displaying first 10 autocorrelations and autocovariances
  acf_values <- acf(wn, lag.max=10, plot=FALSE)$acf[1:10]
  acf_cov <- acf(wn, type="covariance", lag.max=10, plot=FALSE)$acf[1:10]
  
  # Printing autocorrelation and autocovariance values
  print(paste("Autocorrelations for n =", n, ":"))
  print(acf_values)
  print(paste("Autocovariances for n =", n, ":"))
  print(acf_cov)
}

# Analyzing for different sample sizes
sample_sizes <- c(10, 50, 100, 10000)
for (n in sample_sizes) {
  analyze_white_noise(n)
}
```
### Correlogram Patterns
#As the sample size increases, the correlograms become more stable. The white noise process is better represented, and autocorrelations approach zero. This trend highlights the importance of sample size in estimating autocorrelations.

###Effect of Variance on ACF
#Changing the variance (e.g., generating data from N(0, σ^2 ≠ 1)) will alter the autocovariances, but the ACF (autocorrelation function) will remain unchanged due to its scale-invariant nature.

```{r}
# Generate white noise with variance different from 1
wn_var2 <- rnorm(100, mean=0, sd=2)  # Variance = 4 (σ^2 = 2^2)

# Set up plotting layout
par(mfrow=c(2,1))

# Plot trajectory of white noise with variance = 4
plot(wn_var2, type='l', ylab="White Noise", main="Trajectory of White Noise (σ^2 = 4)")
points(wn_var2)

# Plot ACF (correlogram) for white noise with variance = 4
acf(wn_var2, lag.max=10, plot=TRUE, main="Correlogram (σ^2 = 4)")

# Display first 10 autocorrelations and autocovariances
acf_values_var2 <- acf(wn_var2, lag.max=10, plot=FALSE)$acf[1:10]
acf_cov_var2 <- acf(wn_var2, type="covariance", lag.max=10, plot=FALSE)$acf[1:10]

print("Autocorrelations for σ^2 = 4:")
print(acf_values_var2)
print("Autocovariances for σ^2 = 4:")
print(acf_cov_var2)
```

### Conclusion

#Correlograms Pattern: As the sample size increases, the correlograms stabilize, and the autocorrelations approach zero, accurately depicting white noise behavior.

#Effect of Variance on ACF: The ACF remains the same for different variances, demonstrating its scale invariance, while autocovariances change with variance.

